On the minimum leaf conductance: its role in models of plant water use, and ecological and environmental controls

### 2018-06-02 manuscript.Rmd
update rosana contact and acknowledgment
co-author comments incorporated, order changed again; png removed

\
Remko A. Duursma^1^, Chris J. Blackman^1^, Rosana Lopéz^1^, Nicolas K. Martin-StPaul^2^, Hervé Cochard^3^, Belinda E. Medlyn^1^

@@

```{r include=FALSE}
includefigs <- FALSE
library(png)
library(grid)
library(knitr)
library(pander)

@@

```{r, echo=FALSE}
tab <- read.csv(text="I. Introduction
II. Use of minimum conductance in models
\t1. Models of water-use efficiency
\t2. Models of plant desiccation
\t3. Towards a new model formulation
III. Comparison of various definitions of minimum conductance
IV. Problems with estimating Minimum conductance from regression
V. Cuticular conductance
VI. Contribution of stomata
VII. Nighttime conductance
VIII. Environmental and ecological variation in minimum conductance 
\t1. Contribution of stomata to minimum conductance
\t2. Minimum conductance shows little meaningful interspecific variation
\t3. Acclimation to the environment
IX. Conclusions", header=F)

knitr::kable(tab, col.names="")

@@

During conditions where photosynthesis does not take place, such as severe drought, extreme temperature or low light, it seems advantageous for plants to close stomata and completely halt water loss. However, water loss continues through the cuticle and incompletely closed stomata, together constituting the leaf minimum conductance (g~min~). In this review, we illustrate the role of g~min~ in models of leaf gas exchange, critically evaluate sources of variation in g~min~, and quantitatively compare various methods for its estimation. We pay special attention to the role of the minimum conductance in the Ball-Berry model of stomatal conductance, and caution against the usual regression-based method for its estimation. A literature compilation of g~min~ as measured by weight loss of detached leaves is presented, which shows much variation in this trait that is not clearly related to species groups, climate of origin, or leaf type. Much evidence points to the idea that g~min~ is highly responsive to the growing conditions of the plant, including soil water availability, temperature, and air humidity - as we further demonstrate with two case studies. The synthesis presented here provides guidelines for use of g~min~ in ecosystem models, and points to clear research gaps for this drought tolerance trait.

@@

\
The absolute minimum attainable water loss rate is through the cuticle only, which is typically measured on isolated cuticles of the adaxial (non-stomatal) side of the leaf. Other measurements allow for the estimation of water loss through incompletely closed and broken stomata. Of particular interest because of its simplicity is the method where leaves are detached from the plant, and their weight loss monitored over time. This method (mass loss of detached leaves, MLD) aims to simulate field conditions during severe drought, when water supply to the leaf has practically ceased, however proceeds much quicker (typically 0.5-2 days). This measurement also includes both surfaces of the leaf - not just the stomata-free side. A comprehensive compilation of estimates of g~min~ is currently lacking, and is needed not just to parameterize models, but also to study sources of variation in this overlooked plant trait.

\
In this review, we discuss the role of the minimum conductance in models of plant water use, and critically evaluate sources of variation in this overlooked parameter. Previous reviews of the minimum conductance have largely focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles [@riederer2008, @schuster2017; @kerstiens1996; @fernandez2017]. Other work has focused on a different definition of the minimum conductance (g~0~): the value that should be used in models of plant water use [e.g. @barnard2013; @lombardozzi2017], and includes both g~min~ and a 'stomatal residual' due to the fact that stomata do not completely close during periods of zero photosynthesis [@dekauwe2015]. This discussion often centers around leaf conductance during the night [@lombardozzi2017], but we must also consider appropriate values for the minimum conductance during the day, for example during very high (or low) temperature, transient low light, and extreme drought. Thus there is a disconnect between the dizzying level of understanding of the plant cuticle [@riederer2008], and the very simple assumptions made when using leaf conductance models in global vegetation models.

\
We aim to improve the connection between the physiology of minimum conductance, empirical approaches, and model implementations by synthesizing the state of knowledge. Our goals are to a) demonstrate the need to include a non-zero minimum conductance in models of water-use efficiency and drought responses, and discuss current practices b) compare various definitions of g~min~ quantitively, which have sometimes been previously assumed equal, c) present a new compilation of minimum conductance measurements, test for ecologically meaningful patterns, and demonstrate the large acclimatory potential of this trait to environmental drivers.

**Table 1.** Definitions of minimum conductance.

@@

# II. Use of minimum conductance in models

@@

\
Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, and f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987, see @damour2010 for a list of functions), and g~0~ is the value of g~s~ when A~n~ is zero. The g~1~ parameter is directly related to the water-use efficiency: large values of g~1~ indicate low water-use efficiency. In this paper, we refer to Eq. 1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987). Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the g~0~ parameter has been studied in much less detail. For example, a recent comprehensive review of stomatal conductance models did not mention g~0~ or any similar minimum conductance [@damour2010]. The original description of Eq. 1 as published by @ballberry1987 did not include an intercept term. It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model. In practice, g~0~ is estimated from a regression of Eq. 1 with leaf gas exchange data, but as we will show, this may lead to inaccurate values, as pointed out by @barnard2013 but not explained in detail.

@@

\
Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero. But A~n~ can approach zero for many different reasons, including low light, high temperature, low humidity, and drought, and it is unclear whether the same g~0~ is reached in each of these cases. In the application of the Ball-Berry model, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005]. It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cuti~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata. Photosynthesis also ceases at very high temperature, but stomata do not appear to always close in proportion to this decrease in photosynthesis, if at all [@drake2018; @urban2017].

@@

\
Using the coupled leaf gas exchange model, we show in Fig. 1 how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^). We also demonstrate the effect of g~0~ on modelling the T~leaf~ response. In this simulation, T~leaf~ and VPD are assumed to co-vary with an empirical relationship as used by @duursma2014. Over the entire range of T~leaf~, inclusion of a non-zero g~0~ obviously increases leaf transpiration, but at a slightly higher rate than just due to g~0~. This effect arises because the additional conductance allows slightly higher rates of photosynthesis, which in turn increase g~s~ via Eq. 1.

```{r fig.height=6,echo=FALSE, eval=includefigs}
plotfig("output/figure1.png")
```

\
**Figure 1**. Simulations with a coupled leaf gas exchange model [@duursma2015], demonstrating the effect of inclusion of the g~0~ parameter (Eq. 1) on leaf fluxes. (a) Intrinsic water-use efficiency (A~n~/g~s~) as a function of he photosynthetic photon flux density (PPFD), holding other environmental drivers constant, for three values of g~0~. (b) The same simulations as in (a), but showing the intercellular CO~2~ concentration (C~i~). (c) Leaf transpiration (E~L~) simulations, where vapour pressure deficit (VPD) and air temperature (T~air~) were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions. (d) The same simulations as in (c), but showing C~i~. Note how C~i~ increases at high VPD and T~air~, only when g~0~> 0. For all simulations it is assumed that T~leaf~ is equal to T~air~, and ignoring the differential permeability of the cuticle to CO~2~ and H~2~O [@hanson2016].

@@

\
When plants are sufficiently water-stressed so that stomata are mostly closed, water loss still continues at a rate determined by the minimum conductance. Thus models that aim to predict when plants desiccate and die must include a minimum conductance term. A classic study by @pisekwinkler1953 already calculated the length of time needed to desiccate leaves to some critical low water content, given the minimum transpiration rate, and the saturated water content of the leaves. Based on that work, @burghardt2006 reported on a direct correlation between g~min~ and the survival time of leaves. @sinclair2000 presented the minimum conductance as a key drought tolerance trait, and used it to as a basis for predicting crop mortality during severe drought. More recently, @gleason2014 and @blackman2016 proposed that embolism resistance along with whole plant capacitance and minimum transpiration rates all contribute to define the time to desiccation. Finally, @martinstpaul2017 demonstrated in a whole plant model of hydraulic failure that g~min~ was one of the key parameters to explain the drop in water potential below the cavitation threshold, because stomata generally close well before this threshold. Relying on the *Sureau* model presented in @martinstpaul2017 we illustrate the critical role of g~min~ in defining the desiccation time (Fig. 2).

```{r fig.height=6, echo=FALSE, eval=includefigs}
plotfig("output/figure2.png")
```

>>>>

\
Remko A. Duursma^1^, Chris J. Blackman^1^, Rosana Lopéz^1,3^, Nicolas K. Martin-StPaul^2^, Hervé Cochard^3^, Belinda E. Medlyn^1^

@@

```{r include=FALSE}
library(knitr)
library(pander)

@@

```{r, echo=FALSE}
tab <- read.csv(text=

    "I. Introduction                                                        
    II. Comparison of various definitions of minimum conductance           
    III. Cuticular conductance                                             
    IV. Contribution of stomata                                            
    V. Environmental and ecological variation in minimum conductance       
    \t1. Minimum conductance shows little meaningful interspecific variation
    \t2. Acclimation to the environment                                     
    VI. Use of minimum conductance in models                               
    \t1. Models of water-use efficiency                                     
    \t2. Models of plant desiccation                                        
    \t3. Problems with estimating g~0~ from regression                      
    \t4. Nighttime conductance                                              
    \t5. Towards a new model formulation                                    
    VII. Conclusions",
    header=F)
knitr::kable(tab, col.names="")

@@

During conditions where photosynthesis does not take place, such as severe drought, extreme temperature or low light, it seems advantageous for plants to close stomata and completely halt water loss. However, water loss continues through the cuticle and incompletely closed stomata, together constituting the leaf minimum conductance (g~min~). In this review, we critically evaluate sources of variation in g~min~, quantitatively compare various methods for its estimation, and illustrate the role of g~min~ in models of leaf gas exchange. A literature compilation of g~min~ as measured by weight loss of detached leaves is presented, which shows much variation in this trait that is not clearly related to species groups, climate of origin, or leaf type. Much evidence points to the idea that g~min~ is highly responsive to the growing conditions of the plant, including soil water availability, temperature, and air humidity - as we further demonstrate with two case studies. We pay special attention to the role of the minimum conductance in the Ball-Berry model of stomatal conductance, and caution against the usual regression-based method for its estimation. The synthesis presented here provides guidelines for use of g~min~ in ecosystem models, and points to clear research gaps for this drought tolerance trait.

@@

\
The absolute minimum attainable water loss rate is through the cuticle only, which is typically measured on isolated cuticles of the adaxial (non-stomatal) side of the leaf. Other measurements allow for the estimation of water loss through incompletely closed and broken stomata. Of particular interest because of its simplicity is the method where leaves are detached from the plant, their weight loss monitored over time, and expressed as the minimum conductance (g~min~). This method (mass loss of detached leaves, MLD) aims to simulate field conditions during severe drought, when water supply to the leaf has practically ceased, although it proceeds much more quickly (typically 0.5-2 days, compared to weeks or even months in field conditions). This measurement also includes both surfaces of the leaf - not just the adaxial side. A comprehensive compilation of estimates of g~min~ is currently lacking, and is needed not just to parameterize models, but also to study sources of variation in this overlooked plant trait.

\
In this review, we discuss the role of the minimum conductance in models of plant water use, and critically evaluate sources of variation in this parameter. Previous reviews of the minimum conductance have largely focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles [@riederer2008; @schuster2017; @kerstiens1996; @fernandez2017]. Other work has focused on a different definition of the minimum conductance (g~0~): the value that should be used in models of plant water use [e.g. @barnard2013; @lombardozzi2017], which includes both g~min~ and a 'stomatal residual' due to the fact that stomata do not completely close during periods of zero photosynthesis [@dekauwe2015]. This discussion often centers around leaf conductance during the night [@lombardozzi2017], but we must also consider appropriate values for the minimum conductance during the day, for example during very high (or low) temperature, transient low light, and extreme drought. Thus there is a disconnect between the detailed understanding of the plant cuticle [@riederer2008], and the very simple assumptions made when using leaf conductance models in global vegetation models.

\
We aim to improve the connection between the physiology of minimum conductance, empirical approaches, and model implementations by synthesizing the state of knowledge. Our goals are to a) quantitively compare various definitions of g~min~, which have sometimes been assumed equal, b) present a new compilation of minimum conductance measurements, which we use to test for ecologically meaningful patterns, and demonstrate the large acclimatory potential of this trait to environmental drivers, and c) demonstrate the need to include a non-zero minimum conductance in models of water-use efficiency and drought responses.

\
**Table 1.** Definitions of minimum conductance.

@@

# II. Comparison of various definitions of minimum conductance

\
Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories. The absolute minimum conductance attained by leaves is that through the cuticle only. Measurements of cuticular conductance (g~cuti~) are typically made on isolated non-stomatal cuticles [@riederer2001], via special gas exchange techniques [@boyer1997], or mass loss of detached leaves by sealing the side with stomata [@kerstiens1996]. We further discuss some important aspects of cuticular conductance in Section III.

\
We contrasted g~cuti~ and g~min~ with three estimates of stomatal conductance when photosynthesis rates are low or zero during non-drought conditions: night-time conductance (g~dark~) (further discussed in Section VI.4), conductance at low photosynthetically active radiation (PAR) (0 - 40 $\mu$ mol m^-2^ s^-1^) during the day, and conductance during conditions when photosynthesis rates are very low (< 1 $\mu$ mol m^-2^ s^-1^) (but excluding low PAR and drought).

\
We quantitatively compared each of the different conductances by synthesizing existing data. Measurements of conductance of isolated, non-stomatal cuticles (g~cuti~) were taken from the compilations by @kerstiens1996 and @schuster2017 (both also include g~min~ data but these were not used since they included no metadata). Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see Methods S1 for description and references). For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources (see Methods S1 for full details of the compilation). The database includes a total of 221 species (136 Angiosperm, 49 Gymnosperm, 1 Pteridophyte), from 57 taxonomic families, with woody species making up roughly 2/3 of the data.

\
All conductance values were converted to per unit projected surface area, allowing direct comparison to stomatal conductance data which are typically presented in these units. Estimates of g~dark~ were taken from @lombardozzi2017, who compiled measurements of ca. 150 species during the night (we selected only gas exchange based measurements from their database). Daytime values (g~low PAR~ and g~low A~) were estimated from an update to the @lin2015 database by taking the appropriate subsets, and averaging the g~s~ values by species within study. All data and code to reproduce the database and analyses are available online (see Acknowledgments).

\
The results of the quantitative comparison are presented in Fig. 1. From this compilation, a few striking differences among estimates are apparent. Cuticular conductance averaged 0.45 mmol m^-2^ s^-1^, ten-fold lower compared to g~min~ (4.9 mmol m^-2^ s^-1^). This result is in line with the methods comparison of @kerstiens1996. More recently, @schuster2017 argued, based on data for 8 species, that g~min~ from MLD is comparable to g~cuti~. However in the same study, a literature compilation of hundreds of values showed that g~cuti~ was on 10-fold lower compared to g~min~, though this finding was not reported by @schuster2017 [see also @fernandez2017], and instead g~min~ and g~cuti~ were pooled in their analysis. Nonetheless, the comparison between g~cuti~ and g~min~ should be viewed with some caution as @kerstiens1996 argued that measurements of g~cuti~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions. @boyer1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases.

\
Conductance in the dark (g~dark~) (mean = 41.4 mmol m^-2^ s^-1^) was on average eight-fold higher than g~min~. It is no new observation that stomata do not close fully in the dark (further discussed in Section VI.4). Previous studies have directly compared g~dark~ and g~min~ on the same species, and generally concluded that g~dark~ is much higher than g~min~ [@hygen1954; @koerner1994; @walden-coleman2013], but an exception is @cavender-bares2007 who found the two rates to be similar in oaks. In turn, g~dark~ could not be differentiated from stomatal conductance measurements during daytime in very low light (mostly just after dawn, or before sunset) (g~low PAR~), or during conditions when photosynthetic rates were very low (due to very high temperature, VPD, or other factors) (g~low A~). These comparisons demonstrate that g~s~ during non-drought conditions when photosynthetic rates are zero or negligible are much higher than the minimum reached in simulated drought conditions (g~min~).

\
**Figure 1**. Comparison of various estimates of the (presumed) minimum conductance. Error bars are 95\% confidence intervals. Grey dots are the original data (but a few data points occur outside the figure range). Different letters denote significant differences (at $\alpha$ = 0.05). g~cuti~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~low PAR~ leaf conductance during low light (photosynthetically active radiation, PAR); g~low A~ leaf conductance during periods of very low photosynthesis. See text for data sources and methods.

# III. Cuticular conductance

\
A review of cuticular transport mechanisms, biochemical composition, and formation of cuticular waxes is well outside the scope of this review, as these topics have been well described elsewhere [@kerstiens1996; @kerstiens2006; @shepherd2006; @schreiber1996; @schreiber2001; @schuster2016b]. However, a few key points should be summarized as they are relevant to the current discussion, in particular when we aim to interpret variation in literature values of g~min~ (Section V).

\
Although we use the term cuticular conductance freely, this transport pathway does not represent a true conductance since water does not diffuse as a gas through the cuticle. Instead, it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle [@kerstiens1996; @riederer2001; @schreiber1996]. The main barrier to diffusion is actually a very thin layer of wax at the leaf surface. Because most of the resistance is located in such a thin layer, g~cuti~ does not correlate with the thickness of the cuticle [@priestley1943; @riederer2001; @anfodillo2002; @schuster2016b]. Nonetheless there is considerable variation in cuticle thickness among plant species [@schuster2016a], along altitudinal transects [@delucia1984], and even with increasing height in the canopy of very tall trees [@woodruff2010]. If the cuticle thickness does not directly affect its conductance, what is the value of increasing the cuticle?

# IV. Contribution of stomata

The minimum leaf conductance includes two pathways: across the cuticle, and through the (potentially incompletely closed) stomata. Only a few studies have directly quantified the stomatal component of minimum conductance, in contrast to the wealth of information on the cuticular component. A detailed study of *Hedera helix*  concluded that 35% of water loss occurred across the stomatal pores (though closed), and 65% across the non-stomatal part of the cuticle, despite the fact that the stomata presumably covered only a small fraction of the leaf [@santrucek2004]. For ten *Sorghum* genotypes, @muchow1989 reported a strong positive correlation between stomatal density and g~min~, suggesting leaky stomata contribute substantially to g~min~. For seven conifer species, @brodribb2014 reported that 50 - 94% of water loss of detached leaves originated from the stomatal side, concluding that stomata must be very leaky, and likely incompletely closed. However @santrucek2004 found that the cuticle of the stomatal side (but excluding stomata itself) in *Hedera helix* was many times more permeable than the non-stomatal side, thus providing another potential explanation for this difference. Either way, water loss rates of intact leaves cannot be simply explained from the permeability of isolated non-stomatal cuticles, and the contribution of leaky stomata, and potentially more permeable cuticle on the abaxial side, likely varies among species.

\
One key question is whether incomplete stomatal closure is under the plant's control, or whether it is an unavoidable consequence of imperfect stomata, damage, blocking by particles, etc. For example, endophytic fungi - which commonly colonize plant leaves, have been shown to prevent stomatal closure and greatly increase water use  [@arnold2007]. As we discuss in the next section, there is ample evidence to suggest that stomata not only stay open in the dark, but that plants actively control stomatal conductance during the night. During severe drought, there is less evidence for such active control, and it seems likely incomplete closure is not under the plant's control as there is no obvious reason to keep stomata open. Some conditions (especially rapid drying conditions) may lead to excessive drying of the epidermis, which can physically pull apart the stomata ['mechanical advantage', @buckley2005].

# V. Environmental and ecological variation in minimum conductance

\
In the following sections, we review the quite substantial literature on g~min~ measured with the mass loss of detached leaves (MLD) technique, stretching back to the 1930s [@pisekberger1938], focussing on environmental and ecological determinants. We also further analyze the literature compilation of g~min~ (Fig. 2), and separately analyze crop species and their genotypic variation. Comparing all available data, we did not find significant relationships between g~min~ and climate of origin, nor were there meaningful relationships with other traits (see Methods S1). We thus found it difficult to explain the variation in g~min~ among species. A simple breakdown by taxonomic order (Fig. 2b) revealed that grasses (*Poales*) have a higher g~min~ compared to other orders, and conifers (especially *Pinales*) tended towards the lower range of values (but were only significantly different from *Poales*).

\
Very few studies have found meaningful correlations between g~min~ and environmental factors or ecological categories. Based on their own compilation of the literature, @schuster2017 concluded no significant differences in g~min~ or g~cuti~ (analyzed together) by plant growth form. One notable exception is @brodribb2014, who showed a correlation between rainfall at species origin (of the driest quarter), but these authors used a gas-exchange approach (over many weeks), not MLD. This correlation did not hold across the species in our database. Of particular note, *Eucalyptus* species (n = 11, included in *Myrtales*) - all measured on the driest continent of Australia - have g~min~ values slightly higher than the average (Fig. 2b).

\
In crop science, g~min~ has long been identified as a key drought tolerance trait [@sinclair1986]. A number of studies have targeted g~min~ as a key trait for breeding more drought-tolerant crops, leading to comparisons of g~min~ across genotypes grown in the same conditions. For example, @james2008 compared g~min~ in 58 soybean (*Glycine max*) genotypes, and found > 2-fold variation that could not be easily explained by other traits. In Fig. 2c, we have compiled a number of studies in crops, demonstrating not just variation in g~min~ among crops, but also the wide range in g~min~ among genotypes. Again, it is striking that the wide variation in g~min~ cannot be easily explained by variation in leaf or other traits or chemical and structural components of the cuticle [@bengtson1978; @james2008; @saito2010], suggesting a significant role of incomplete stomatal closure.

\
**Figure 2**. Analysis of a literature compilation of g~min~ estimates, as measured with mass loss of detached leaves. (a) Histogram (probability density) of all estimates (after averaging by species, n=`r nrow(gmindat)`), with a log-normal distribution curve (mean = `r round(exp(mean(log(gmindat$gmin))),2)`, sd = `r round(exp(sd(log(gmindat$gmin))),2)`). (b) g~min~ averaged by phylogenetic order (including only the top ten orders in the database). Bars are 95% confidence intervals. Numbers above the figure refer to the number of species. Different letters denote significant differences (at $\alpha$ = 0.05, adjusted for multiple comparisons), and grey symbols are species-level data. (c) g~min~ estimates for crops only, averaged by genotype. Bars denote the range, illustrating the wide range in g~min~ among genotypes for a particular crop species. Numbers above the figure refer to the number of genotypes included.

## 2. Acclimation to the environment

\
While there is clearly considerable variation in g~min~ among species, a number of lines of evidence suggest that g~min~ also has great potential for plasticity. Here we summarize the literature on the acclimation of g~min~ to drought conditions, to changes in temperature and humidity, and the change in g~min~ with leaf age, altitude and other factors. This discussion is directly relevant to the use of g~min~ in models, because if the degree of plasticity is large, it complicates model parameterization.

\
There is a general tendency for a decreased g~min~ in plants acclimated to drought stress [@james2008]. The magnitude of the decrease of g~min~ with acclimation to drought stress varied from -4 to -70% (across 10 studies, see Table S2), with a typical decrease in the order of 30-40%. In each of the studies summarized, plants were grown in well-watered or droughted conditions, and in one case a difference in g~min~ was demonstrated after just four days of drought exposure [@bengtson1978]. We also demonstrate drought acclimation via a case study on 11 *Hakea* species grown in two watering treatments (Fig. 3). All 11 species showed a decrease in g~min~ in the drought treatment (see Methods S3 for experimental details). This significant change in g~min~ with drought acclimation is likely an important component of the overall drought hardening of plants.

\
A number of studies have attempted to attribute the drought-induced acclimation in g~min~ to a change in the chemical composition of the cuticle. @bengtson1978 reported lower g~min~ in six oat varieties in response to drought, and an increase in the amount of cuticular waxes, but could not find a relationship between the two and the response was highly genotype-specific. However, @premachandra1992 found that epicuticular wax load increased on leaves of non-irrigated Sorghum cultivars and was positively correlated with cuticular conductance and cell membrane stability. @mackova2013 found that adding ABA (simulating drought stress) to *Lepidium sativum* increased the chain length of cuticular waxes (but not total amounts). @bi2017 also reported that drought caused a change in cuticular wax production and chemical composition, but again in a highly genotype-specific way. Thus it can be concluded that the cuticle indeed changes substantially after a change in plant available water, but in a complex, species-specific manner that is yet to be connected directly to changes in g~min~.

\
**Figure 3**. Minimum conductance (g~min~) measured by mass loss of detached leaves on a variety *Hakea* species, a genus native to Australia. Plants were grown in containers in a grow house, and supplied with ample water or subjected to long-term (8 months) mild drought stress. Bars are labelled by leaf form : broadleaf (B) or needle-like (N), and ordered by g~min~ in the  well-watered treatment. For droughted plants, g~min~ was higher for species with needle-like leaves (Wilcoxon test, p `r pval(wilcox.test(gmin.d ~ leaf.form, data=lopw)$p.value)`)), but not for well-watered plants (p `r pval(wilcox.test(gmin.w ~ leaf.form, data=lopw)$p.value)`).

\
The response of g~min~ to temperature is more complex; it shows both a response to instantaneous changes in temperature, as well as an acclimatory response to growth temperature. @riederer2001 and @schuster2016b have both demonstrated a steep non-linear instantaneous response of g~cuti~ to temperature, with the response becoming especially steep at higher temperatures (> 40C). In *Eucalyptus haemastoma*, the response of g~min~ to temperature was so steep, that the proportion of cuticular to total transpiration increased from 2-3% at 20°C, to 40% at 38°C [@eamus2008]. The mechanism of the instantaneous temperature response is complex, highly species-specific, and we refer to @schuster2016b for a detailed investigation. The rapid increase in g~min~ at high temperature may well be a crucial component of the ability of plants to tolerate heat waves [@drake2018]. In support of the link between g~min~ and heat tolerance, @schuster2016a reported on a negative relationship between thermal tolerance and g~min~ across nine species, so that species with improved tolerance to very high temperature had a low g~min~.

\
Less well established is the acclimatory potential of g~min~ to changes in growth temperature. We present a case study on *Eucalyptus parramattensis* grown in whole-tree chambers (see Methods S3 for experimental details). The chambers either tracked ambient conditions, or were subjected to a +3°C warming treatment. After several months of growth in the treatments, g~min~ was measured at various temperatures, ranging from 17.5 to 27.5°C. We found a 56% decrease in g~min~ in the elevated temperature trees, but there was no clear pattern with measurement temperature (Fig. 4). This decrease in g~min~ is consistent with the drought response, because again g~min~ is reduced in leaves that are subjected to conditions that increase evaporative demand [vapour pressure deficit was higher in the elevated temperature treatment, see @drake2018]. The direction of this response is consistent with @duarte2016, who reported lower g~0~ and g~dark~ in a heat wave treatment in *Pseudotsuga menziesii*, which persisted for some time. Responses across temperature gradients are more complex. In particular, the change in g~min~ with altitude has been well studied, and it is commonly  reported that g~min~ increases with altitude [e.g. @delucia1984; @herrick1991; @anfodillo2002]. @fernandez2017 discuss the literature on altitude responses in detail and argue that the short growing seasons at high altitude are insufficient for complete maturation of leaf cuticles.

\
Finally we mention the striking effect of leaf age on g~min~. We have summarized five studies in Fig. S3, all of which reported an increased g~min~ for older leaves for the majority of species studied. The data reported by @jordan2007 are particularly impressive, as g~min~ in the woody shrub *Agastachys odorata* gradually increased with leaf age up to ca. 10 years. If this effect is due to properties of the cuticle alone, a possible explanation for the increase in g~min~ is the continued exposure to wind, rain, and abrasives, which have been shown to damage the cuticle and increase its conductance (see discussion in Section III). Another possibility is that the contribution of stomata to g~min~ increases with leaf age, as reported by @jordan2007. In plants that maintain several cohorts of leaves, finding an appropriate value of g~min~ for use in models will have to take into account the leaf age effect.

\
**Figure 4**. Minimum conductance (g~min~) measured with mass loss of detached leaves on *Eucalyptus parramattensis* grown in whole-tree chambers. Trees were grown following ambient conditions, or in an elevated temperature treatment (+3 °C) [see @drake2018]. Measurements of g~min~ were carried out at five different temperatures on replicate leaves. There was no consistent effect of measurement temperature on g~min~, but leaves of trees grown in elevated temperature showed on average a 45% reduction in g~min~ (p < 0.01, linear mixed-effects model).

# VI. Use of minimum conductance in models

@@

\
Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987, see @damour2010 for a list of functions), and g~0~ is the value of g~s~ when A~n~ is zero. The g~1~ parameter is directly related to the water-use efficiency: large values of g~1~ indicate low water-use efficiency. In this paper, we refer to Eq. 1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987). Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the g~0~ parameter has been studied in much less detail. For example, a recent comprehensive review of stomatal conductance models did not mention g~0~ or any similar minimum conductance [@damour2010]. The original description of Eq. 1 as published by @ballberry1987 did not include an intercept term. It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model. In practice, g~0~ is usually estimated from a regression of Eq. 1 with leaf gas exchange data, but as we will show, this approach may lead to inaccurate values, as pointed out by @barnard2013 but not explained in detail.

@@

\
Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero. However, A~n~ can approach zero for many different reasons, including low light, high temperature, low humidity, and drought, and it is unclear whether the same g~0~ is reached in each of these cases. In the application of the Ball-Berry model, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005]. An alternative common assumption is that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cuti~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata. Photosynthesis also ceases at very high temperature, but stomata do not appear to always close in proportion to this decrease in photosynthesis, if at all [@drake2018; @urban2017].

@@

\
Using the coupled leaf gas exchange model, we show in Fig. 5 how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^). We also demonstrate the effect of g~0~ on modelling the T~leaf~ response. In this simulation, T~leaf~ and VPD are assumed to co-vary with an empirical relationship as used by @duursma2014. Over the entire range of T~leaf~, inclusion of a non-zero g~0~ obviously increases leaf transpiration, but at a slightly higher rate than just due to g~0~. This effect arises because the additional conductance allows slightly higher rates of photosynthesis, which in turn increase g~s~ via Eq. 1.

\
**Figure 5**. Simulations with a coupled leaf gas exchange model [@duursma2015], demonstrating the effect of inclusion of the g~0~ parameter (Eq. 1) on leaf fluxes. (a) Intrinsic water-use efficiency (A~n~/g~s~) as a function of he photosynthetic photon flux density (PPFD), holding other environmental drivers constant, for three values of g~0~. (b) The same simulations as in (a), but showing the intercellular CO~2~ concentration (C~i~). (c) Leaf transpiration (E~L~) simulations, where vapour pressure deficit (VPD) and air temperature (T~air~) were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions. (d) The same simulations as in (c), but showing C~i~. Note how C~i~ increases at high VPD and T~air~, only when g~0~> 0. For all simulations it is assumed that T~leaf~ is equal to T~air~, and ignoring the differential permeability of the cuticle to CO~2~ and H~2~O [@hanson2016].

@@

\
When plants are sufficiently water-stressed so that stomata are mostly closed, water loss still continues at a rate determined by the minimum conductance. Thus models that aim to predict when plants desiccate and die must include a minimum conductance term. A classic study by @pisekwinkler1953 calculated the length of time needed to desiccate leaves to some critical low water content, given the minimum transpiration rate, and the saturated water content of the leaves. Based on that work, @burghardt2006 reported on a direct correlation between g~min~ and the survival time of leaves. @sinclair2000 presented the minimum conductance as a key drought tolerance trait, and used it to as a basis for predicting crop mortality during severe drought. More recently, @gleason2014 and @blackman2016 proposed that embolism resistance along with whole plant capacitance and minimum transpiration rates all contribute to define the time to desiccation. Building on this work, @martinstpaul2017 demonstrated in a whole plant model of hydraulic failure that g~min~ was one of the key parameters to explain the drop in water potential below the cavitation threshold, because stomata generally close well before this threshold. Applying the *Sureau* model presented in @martinstpaul2017 we illustrate the critical role of g~min~ in defining the desiccation time (Fig. 6).

\
**Figure 6**. Simulations with the *Sureau* model demonstrating the effect of g~min~ on desiccation tolerance of plants. The *Sureau* model simulates water transport in the soil-plant-atmosphere continuum, and includes a detailed representation of capacitance in stem and leaf tissues. (a) Soil relative extractable water (REW; 1 = field capacity, 0 = permanent wilting point) for the two simulations, using a minimum conductance (g~min~) of 2 or 4 mmol m^-2^ s^-1^ - all other parameters were equal. (b) Water potential in the soil and leaf as the dry-down progresses. (c) Progression of percent loss conductivity (PLC) of the xylem. Dashed line is at a PLC of 88%, indicating possible mortality.

## 3. Problems with estimating g~0~ from regression
